---
title: "compound possion with logistic flash"
author: "Wei Wang"
date: 2017-03-08
output: html_document
---

<!-- The file analysis/chunks.R contains chunks that define default settings
shared across the workflowr files. -->
```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

<!-- Update knitr chunk options -->
```{r knitr-opts-chunk, include=FALSE}
```

<!-- Insert the date the file was last updated -->
```{r last-updated, echo=FALSE, results='asis'}
```

<!-- Insert the code version (Git commit SHA1) if Git repository exists and R
 package git2r is installed -->
```{r code-version, echo=FALSE, results='asis'}
```

<!-- Add your analysis here -->

## Compound Possion model with logistic flash

### model

\begin{eqnarray}
Y_{ij}|\lambda_{ij},p_{ij} &=& poisson(\lambda_{ij} p_{ij}) \\
logit(p_{ij}) &=& L_i F_j \\
\lambda_{ij} &\sim& g_{ash}(\cdot)\\ 
L_i &\sim& g_{ash}(\cdot) \\
F_j &\sim& g_{ash}(\cdot)
\end{eqnarray}

#### write the model in compound possion distribution

\begin{eqnarray}
Y_{ij}|n_{ij},p_{ij} &=& Bin(n_{ij},p_{ij}) \\
logit(p_{ij}) &=& L_i F_j \\
n_{ij} &\sim& Poisson(\lambda_{ij})\\
\lambda_{ij} &\sim& g_{ash}(\cdot)\\ 
L_i &\sim& g_{ash}(\cdot) \\
F_j &\sim& g_{ash}(\cdot)
\end{eqnarray}



### inference

#### 1  $p_{ij}|Y_{ij},n_{ij}$

we can apply logistic flash on $p_{ij}$ given $n_{ij}$, since given $n_{ij}$ the problem is a logistic flash model:

\begin{eqnarray}
Y_{ij}|n_{ij},p_{ij} &=& Bin(n_{ij},p_{ij}) \\
logit(p_{ij}) &=& L_i F_j \\
L_i &\sim& g_{ash}(\cdot) \\
F_j &\sim& g_{ash}(\cdot)
\end{eqnarray}

Click on this [details](third-analysis.html) to see my result.

#### 2  $n_{ij}|p_{ij},Y_{ij},\lambda_{ij}$

The marginalization disctribution of $n_{ij}$ given $p_{ij}$ is 

\begin{eqnarray}
n_{ij} \sim Y_{ij} + poisson(\lambda_{ij}(1-p_{ij}))
\end{eqnarray}
this can be calculated as follows
\begin{eqnarray}
p(n_{ij}|p_{ij},\lambda_{ij},Y_{ij}) &\propto& p_{ij}^{Y_{ij}}(1-p_{ij})^{n_{ij} - Y_{ij}}\frac{n_{ij}!}{Y_{ij}!(n_{ij} - Y_{ij})!} \frac{\lambda_{ij}^{n_{ij}} e^{-\lambda_{ij}}}{n_{ij}!}\\
&\propto&  \frac{(\lambda_{ij}(1-p_{ij}))^{(n_{ij}-Y_{ij})} e^{-(\lambda_{ij}(1-p_{ij}))}}{(n_{ij}-Y_{ij})!}
\end{eqnarray}

#### 3  $\lambda_{ij}|p_{ij},Y_{ij}$

This is the model for ash on poisson likelihood:

\begin{eqnarray}
Y_{ij} &\sim& poisson(\lambda_{ij}p_{ij})\\
\lambda_{ij} &\sim& g_{ash}(\cdot)
\end{eqnarray}



## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```
