---
title: "Flash on HumanTungiPSC"
author: "wei wang"
date: 2017-02-14
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

**Last updated:** `r Sys.Date()`

**Code version:** `r workflowr::extract_commit(".", 1)$sha1`

```{r,warning=FALSE,cache=TRUE,eval=TRUE,message=FALSE}
# Biobase is for accessing expressionSet objects
library(Biobase)

# Load Buettner data
devtools::install_github("jhsiao999/singleCellRNASeqHumanTungiPSC")
library(singleCellRNASeqHumanTungiPSC)

# Extract expression data
eset <- get(data("HumanTungiPSC"))

# Access the count matrix
counts <- exprs(eset)
dim(counts)

Binary_count = 1 * (counts != 0)
Boolean_count = 2*( Binary_count - 1/2)
# Access cell sample information 
#pData(eset)
#cluster <- pData(eset)$Cluster
#table(cluster)
barlabels = pData(HumanTungiPSC)
total_counts = apply(counts,2,sum)

```

## FLASH on the log CPM data

### total log CPM data without imputation

```{r,warning=FALSE,cache=TRUE,eval=TRUE,message=FALSE}
total_logcpm = readRDS("~/HG/LogisticFlash_workflow/data/singelcell/HumanTungiPSC/logcpm/bflash_var_col.rds")
par(mfrow = c(3,3),mar = c(5,4,4,2) - 1.9)
for(i in 1:16) barplot(total_logcpm$l[,i])
pve = sapply(seq(1,26),function(x){sum((total_logcpm$l[,x] %*% t(total_logcpm$f[,x]))^2)})
plot(pve)
plot(log(pve),main = "log")
```

### missing pattern

```{r,warning=FALSE,cache=TRUE,eval=TRUE,message=FALSE}
MissPattern = readRDS("~/HG/LogisticFlash/data/MissPatternHumanTungiPSC_K20.rds")
par(mfrow = c(3,3),mar = c(5,4,4,2) - 1.9)
for( i in 1:8) barplot(MissPattern$f[,i],main = paste("factor",i))
pve = sapply(seq(1,8),function(x){sum((MissPattern$l[,x] %*% t(MissPattern$f[,x]))^2)})
# plot(pve)
plot(log(pve),main = "log")
```


```{r}
non_miss_logcpm = readRDS("~/HG/LogisticFlash_workflow/data/singelcell/HumanTungiPSC/logcpm/bflash_non_miss_var_col.rds")
par(mfrow = c(3,3),mar = c(5,4,4,2) - 1.9)
for(i in 1:16) barplot(non_miss_logcpm$l[,i])
pve = sapply(seq(1,30),function(x){sum((non_miss_logcpm$l[,x] %*% t(non_miss_logcpm$f[,x]))^2)})
plot(pve)
plot(log(pve),main = "log")
```


### log cpm with impute

```{r,warning=FALSE,cache=TRUE,eval=TRUE,message=FALSE}
logcpm_imput = readRDS("~/HG/LogisticFlash_workflow/data/singelcell/HumanTungiPSC/logcpm/gflash_var_col_imput.rds")
par(mfrow = c(3,3),mar = c(5,4,4,2) - 1.9)
for(i in 1:19) barplot(logcpm_imput$l[,i])
pve = sapply(seq(1,19),function(x){sum((logcpm_imput$l[,x] %*% t(logcpm_imput$f[,x]))^2)})
plot(pve)
plot(log(pve),main = "log")
```

## compare the impute and non-impute methods

I find the imputation from flash have small difference from the logcpm from voom.

```{r}
par(mfrow = c(3,3),mar = c(5,4,4,2) - 1.9)
for(i in 1:9) plot(total_logcpm$l[,i],logcpm_imput$l[,i])
```

does the imputaion help or not?

```{r}
par(mfrow = c(2,3),mar = c(5,4,4,2) - 1.9)
library(ggplot2)
plot_df = data.frame(total_logcpm$l[,1:8])
colnames(plot_df) = c("f1","f2","f3","f4","f5","f6","f7","f8")
plot_df$total_count = total_counts
p1 <- ggplot(plot_df, aes(total_count, f4, label = (as.character(factor(barlabels$individual )) )), parse = FALSE)
p1 + geom_label(aes(fill = factor(barlabels$batch)), colour = "white", fontface = "bold")
p1 <- ggplot(plot_df, aes(total_count, f6, label = (as.character(factor(barlabels$individual )) )), parse = FALSE)
p1 + geom_label(aes(fill = factor(barlabels$batch)), colour = "white", fontface = "bold")
p1 <- ggplot(plot_df, aes(total_count, f7, label = (as.character(factor(barlabels$individual )) )), parse = FALSE)
p1 + geom_label(aes(fill = factor(barlabels$batch)), colour = "white", fontface = "bold")
```

```{r}
par(mfrow = c(2,3),mar = c(5,4,4,2) - 1.9)
library(ggplot2)
plot_df = data.frame(logcpm_imput$l[,1:8])
colnames(plot_df) = c("f1","f2","f3","f4","f5","f6","f7","f8")
plot_df$total_count = total_counts
p1 <- ggplot(plot_df, aes(total_count, f4, label = (as.character(factor(barlabels$individual )) )), parse = FALSE)
p1 + geom_label(aes(fill = factor(barlabels$batch)), colour = "white", fontface = "bold")
p1 <- ggplot(plot_df, aes(total_count, f6, label = (as.character(factor(barlabels$individual )) )), parse = FALSE)
p1 + geom_label(aes(fill = factor(barlabels$batch)), colour = "white", fontface = "bold")
p1 <- ggplot(plot_df, aes(total_count, f7, label = (as.character(factor(barlabels$individual )) )), parse = FALSE)
p1 + geom_label(aes(fill = factor(barlabels$batch)), colour = "white", fontface = "bold")
```

## what missing patterns learns?


```{r}
library(ggplot2)
plot_df = data.frame(logcpm_imput$l[,1:8])
colnames(plot_df) = c("f1","f2","f3","f4","f5","f6","f7","f8")
plot_df$total_count = total_counts
p1 <- ggplot(plot_df, aes(total_count, f2, label = (as.character(factor(barlabels$individual )) )), parse = FALSE)
p1 + geom_label(aes(fill = factor(barlabels$individual)), colour = "white", fontface = "bold")
p1 <- ggplot(plot_df, aes(total_count, f3, label = (as.character(factor(barlabels$individual )) )), parse = FALSE)
p1 + geom_label(aes(fill = factor(barlabels$individual)), colour = "white", fontface = "bold")

```


```{r}
par(mfrow = c(3,3),mar = c(5,4,4,2) - 1.9)
which_ercc <- grep("ERCC", rownames(counts))
plot(total_logcpm$f[,2],MissPattern$l[-which_ercc,2])
plot(total_logcpm$f[,3],MissPattern$l[-which_ercc,3])
plot(total_logcpm$f[,4],MissPattern$l[-which_ercc,4])
plot(logcpm_imput$f[,2],MissPattern$l[-which_ercc,2])
plot(logcpm_imput$f[,3],MissPattern$l[-which_ercc,3])
plot(logcpm_imput$f[,4],MissPattern$l[-which_ercc,4])
plot(total_logcpm$f[,2],logcpm_imput$f[,2])
plot(total_logcpm$f[,3],logcpm_imput$f[,3])
plot(total_logcpm$f[,4],logcpm_imput$f[,4])
cor(total_logcpm$f[,2],MissPattern$l[-which_ercc,2])
cor(total_logcpm$f[,3],MissPattern$l[-which_ercc,3])
cor(total_logcpm$f[,4],MissPattern$l[-which_ercc,4])
cor(logcpm_imput$f[,2],MissPattern$l[-which_ercc,2])
cor(logcpm_imput$f[,3],MissPattern$l[-which_ercc,3])
cor(logcpm_imput$f[,4],MissPattern$l[-which_ercc,4])
cor(total_logcpm$f[,2],logcpm_imput$f[,2])
cor(total_logcpm$f[,3],logcpm_imput$f[,3])
cor(total_logcpm$f[,4],logcpm_imput$f[,4])
```


```{r}
which(logcpm_imput$l[,3]>0)
colnames(counts)[445]
any_other = data.frame(rawcount = as.vector(counts[,c(400:444,446:500)]))
out_lier = data.frame(rawcount = counts[,445])
any_other$name = "others"
out_lier$name = colnames(counts)[445]
hist_compare <- rbind(any_other, out_lier)
ggplot(hist_compare, aes(rawcount, fill = name)) + geom_density(alpha = 0.2)
any_other = data.frame(rawcount = as.vector(counts[,c(437:445)]))
out_lier = data.frame(rawcount = counts[,445])
any_other$name = "others"
out_lier$name = colnames(counts)[445]
hist_compare <- rbind(any_other, out_lier)
ggplot(hist_compare, aes(rawcount, fill = name)) + geom_density(alpha = 0.2)
ks.test(counts[,445],as.vector(counts[,c(437:445)]))
```


```{r}
par(mfrow = c(2,2),mar = c(5,4,4,2) - 1.9)
plot(MissPattern$f[440:450,3])
plot(total_logcpm$l[440:450,3])
plot(logcpm_imput$l[440:450,3])
```

```{r}
# in the same batch same well
apply(counts[,437:445],2,sum)
apply(Binary_count[,440:450],2,sum)
batch_Dr2 = counts[,437:445]
cor(counts[,445],counts[,437])
image(cor(batch_Dr2))
image(cor(Binary_count[,437:445]))
```



## Session Information

```{r session-info}
```
