---
title: "missing pattern"
author: "Wei Wang"
date: "2/8/2017"
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

**Last updated:** `r Sys.Date()`

**Code version:** `r workflowr::extract_commit(".", 1)$sha1`


## data from joyice

Try to find the pattern of the missing position from joyce's data. This a count data with many zeros which are consider as "missing". I want to find if there is a low rank pattern for the missing positions. So I treat the data as Boolean data which are `miss` or `non-miss`. To deal with the Boolean data factor analysis. We can apply Logistic Flash.

```{r,warning=FALSE,cache=TRUE,eval=TRUE,message=FALSE}
library(Biobase)
# Install and load the required library
library(Biobase)
# Load Buettner data
devtools::install_github("jhsiao999/singleCellRNASeqHumanBuettnerTCell")
library(singleCellRNASeqHumanBuettnerTCell)
# Extract expression data
eset <- get(data("HumanBuettnerTCell"))
# The count matrix is 7073 gene by 81 samples
counts <- exprs(eset)
#dim(counts)
# View sample information (these samples were experimentally-determined to be in two functionally-distinct clusters)
#pData(eset)
cluster <- pData(eset)$Cluster
# table(cluster)
Binary_count = 1 * (counts != 0)
Boolean_count = 2*( Binary_count - 1/2)
```


I run Logistic flash (greedy algorithm) with $K=20$ and it only gives me 3 factors.

```{r,warning=FALSE,cache=TRUE,eval=TRUE,message=FALSE}
source('~/HG/LogisticFlash/Rcode/GD_Rfuncrtions.R')
# MissPattern = GL_flash(Boolean_count,K = 20)
# saveRDS(MissPattern,file = "~/HG/LogisticFlash/data/MissPatternHumanBuettner_K20.rds")
MissPattern = readRDS("~/HG/LogisticFlash/data/MissPatternHumanBuettner_K20.rds")
```


## factors: need the name of columns

I think I need the name of each cell type or labels for the cell type.

```{r}
par(mfrow = c(2,2),mar = c(5,4,4,2)-1.9)
barplot(MissPattern$f[,1],main = "factor 1")
barplot(MissPattern$f[,2],main = "factor 1")
barplot(MissPattern$f[,3],main = "factor 1")
```

### Is there missing pattern?

I think so. If there is only one factor which is common factor, I thinks it just stand for the common signal or the common missing position. The factor 2 and factor 3 should contain the information of missing pattern.

## plot using the cluster in the data

**I don't know what is the** `cluster` and I guess this should be some useful label for the data. Here I think the factor 2 and factor 3 contains the information of the missing position, so I plot the factor 2 vs factor 3. There is some pattern in this plot. I use the label beacuse I can't find other labels. :)

### other type of labels?

I think there is some clustering pattern in the plot, but the cluster might be different from the label `cluster` in the data. (I don't know what is `cluster`.) 


```{r}
library(ggplot2)
plot_df_f = data.frame(MissPattern$f)
p <- ggplot(plot_df_f, aes(V3, V2, label = (as.character(cluster) )))
# p + geom_text()
p + geom_label(aes(fill = factor(cluster)), colour = "white", fontface = "bold")

```



## Session Information

```{r session-info}
```
