---
title: "MouseZeiselBrain"
author: "wei wang"
date: 2017-02-20
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

**Last updated:** `r Sys.Date()`

**Code version:** `r workflowr::extract_commit(".", 1)$sha1`

## load the data

there is no ercc in this data set.

```{r}
#devtools::install_github("jhsiao999/singleCellRNASeqMouseZeiselBrain")

library(singleCellRNASeqMouseZeiselBrain)

counts <- exprs(MouseZeiselBrain)

cell_labels <- pData(MouseZeiselBrain)
col_sum = colSums(counts)
row_sum = rowSums(counts)
col_index = which(col_sum != 0)
row_index = which(row_sum != 0)
# exclude the missing values
count_data = counts[row_index,col_index]

dim(count_data)
```

## missing pattern

There are 20 rank in our result, so there might more rank.

```{r}
MissPattern = readRDS("~/HG/LogisticFlash_workflow/data/singelcell/MouseZeiselBrain/MissPatternMouseZeiselBrain_K20.rds")
```


## no structure in age, sex or tissue

```{r}
 par(mfrow = c(3,4), mar = c(5,4,4,2) - 1.9)
for(i in 1:3){
  plot(as.numeric(as.factor(cell_labels$tissue[col_index])),MissPattern$f[,i],main = paste("f",i,": tissue" ))
  plot(as.numeric(as.factor(cell_labels$sex[col_index])),MissPattern$f[,i],main = paste("f",i,": sex"))
  plot(as.numeric(as.factor(cell_labels$age[col_index])),MissPattern$f[,i],main = paste("f",i,": age"))
  plot(as.numeric(as.factor(cell_labels$level1_class[col_index])),MissPattern$f[,i],main = paste("f",i,": level1"))
}
```

## level1_class

```{r}
plot(MissPattern$f[,2],MissPattern$f[,3])
unique(cell_labels$level1_class)
class_level = unique(cell_labels$level1_class)
class_index = list()
for(i in 1:7){class_index[[i]] = which(cell_labels$level1_class == class_level[i])}
for(i in 1:7){
points(MissPattern$f[class_index[[i]],2],MissPattern$f[class_index[[i]],3],col = rainbow(7)[i])
}
```

We can see there are three classes

- class 1: "interneurons" "pyramidal SS" "pyramidal CA1"
- class 2: "oligodendrocytes"
- class 3: "microglia" "endothelial-mural" "astrocytes_ependymal"

## level1: factor 2; factor 3

```{r,echo=FALSE}
library(ggplot2)
# this is the plot index for the count_data
plot_df = data.frame(MissPattern$f[,1:5])
colnames(plot_df) = c("f1","f2","f3","f4","f5")

group_number = cell_labels$level1_class
group_number[which(cell_labels$level1_class == "interneurons" )] = 1
group_number[which(cell_labels$level1_class == "pyramidal SS" )] = 1
group_number[which(cell_labels$level1_class == "pyramidal CA1" )] = 1
group_number[which(cell_labels$level1_class == "oligodendrocytes" )] = 2
group_number[which(cell_labels$level1_class == "microglia" )] = 3
group_number[which(cell_labels$level1_class == "endothelial-mural" )] = 3
group_number[which(cell_labels$level1_class == "astrocytes_ependymal" )] = 3
p1 <- ggplot(plot_df, aes(f2, f3, label = (as.character(factor(group_number)) )), parse = FALSE)
p1 + geom_label(aes(fill = factor(group_number)), colour = "white", fontface = "bold")
```

this structure makes sense based on the paper [check the fig 1 C](http://linnarssonlab.org/pdf/Science%202015a.pdf)

## FLASH

```{r}
bflash = readRDS("~/HG/LogisticFlash_workflow/data/singelcell/MouseZeiselBrain/logcpm/bflash.rds")
bflash_varcol = readRDS("~/HG/LogisticFlash_workflow/data/singelcell/MouseZeiselBrain/logcpm/bflash_var_col.rds")
```


```{r}
par(mfrow = c(3,4), mar = c(5,4,4,2) - 1.9)
for(i in 1:6){
  #par(mfrow = c(2,2), mar = c(5,4,4,2) - 1.9)
  plot(as.numeric(as.factor(cell_labels$tissue[col_index])),bflash$l[,i],main = paste("f",i,": tissue" ))
  plot(as.numeric(as.factor(cell_labels$sex[col_index])),bflash$l[,i],main = paste("f",i,": sex"))
  plot(as.numeric(as.factor(cell_labels$age[col_index])),bflash$l[,i],main = paste("f",i,": age"))
  plot(as.numeric(as.factor(cell_labels$level1_class[col_index])),bflash$l[,i],main = paste("f",i,": level1"))
}
```

```{r}
par(mfrow = c(2,2), mar = c(5,4,4,2) - 1.9)
plot(bflash$l[,2],bflash$l[,3],main = "flash")
unique(cell_labels$level1_class)
class_level = unique(cell_labels$level1_class)
class_index = list()
for(i in 1:7){class_index[[i]] = which(cell_labels$level1_class == class_level[i])}
for(i in 1:7){
points(bflash$l[class_index[[i]],2],bflash$l[class_index[[i]],3],col = rainbow(7)[i])
}

#plot(bflash_varcol$l[,2],bflash_varcol$l[,3],main = "flash column-wise variacne")
# unique(cell_labels$level1_class)
#class_level = unique(cell_labels$level1_class)
#class_index = list()
#for(i in 1:7){class_index[[i]] = which(cell_labels$level1_class == class_level[i])}
#for(i in 1:7){
#points(bflash_varcol$l[class_index[[i]],2],bflash_varcol$l[class_index[[i]],3],col = rainbow(7)[i])
#}

plot(MissPattern$f[,2],MissPattern$f[,3],main = "Logistic flash")
# unique(cell_labels$level1_class)
class_level = unique(cell_labels$level1_class)
class_index = list()
for(i in 1:7){class_index[[i]] = which(cell_labels$level1_class == class_level[i])}
for(i in 1:7){
points(MissPattern$f[class_index[[i]],2],MissPattern$f[class_index[[i]],3],col = rainbow(7)[i])
}
#library(limma)
#voom_ouptut <- voom(count_data, normalization = "none")
#log_cpm = voom_ouptut$E
#Y = t(log_cpm)

# svd_5 = svd(Y,nu = 5,nv = 5)
# saveRDS(svd_5,"~/HG/LogisticFlash_workflow/data/singelcell/MouseZeiselBrain/logcpm/svd_5.rds")
svd_5 = readRDS("~/HG/LogisticFlash_workflow/data/singelcell/MouseZeiselBrain/logcpm/svd_5.rds")
plot(svd_5$u[,2],svd_5$u[,3],main = "PCA")
# unique(cell_labels$level1_class)
class_level = unique(cell_labels$level1_class)
class_index = list()
for(i in 1:7){class_index[[i]] = which(cell_labels$level1_class == class_level[i])}
for(i in 1:7){
points(svd_5$u[class_index[[i]],2],svd_5$u[class_index[[i]],3],col = rainbow(7)[i])
}
tsne_3 = readRDS("~/HG/LogisticFlash_workflow/data/singelcell/MouseZeiselBrain/logcpm/tsne_3.rds")
plot(tsne_3$Y[,2],tsne_3$Y[,3],main = "TSNE")
# unique(cell_labels$level1_class)
class_level = unique(cell_labels$level1_class)
class_index = list()
for(i in 1:7){class_index[[i]] = which(cell_labels$level1_class == class_level[i])}
for(i in 1:7){
points(tsne_3$Y[class_index[[i]],2],tsne_3$Y[class_index[[i]],3],col = rainbow(7)[i])
}
```

## Session Information

```{r session-info}
```
