---
title: "MouseBuettnerESC missing pattern analysis"
author: "wei wang"
date: 2017-02-10
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

**Last updated:** `r Sys.Date()`

**Code version:** `r workflowr::extract_commit(".", 1)$sha1`

## data 

```{r,warning=FALSE,cache=TRUE,eval=TRUE,message=FALSE}
# Biobase is for accessing expressionSet objects
library(Biobase)

# Load Buettner data
devtools::install_github("jhsiao999/singleCellRNASeqMouseBuettnerESC")
library(singleCellRNASeqMouseBuettnerESC)

# Extract expression data
eset <- get(data("MouseBuettnerESC"))

# Access the count matrix
counts <- exprs(eset)
dim(counts)

Binary_count = 1 * (counts != 0)
Boolean_count = 2*( Binary_count - 1/2)
# Access cell sample information 
total_counts = apply(counts,2,sum)
```


```{r,warning=FALSE,cache=TRUE,eval=TRUE,message=FALSE}
source('~/HG/LogisticFlash/Rcode/GD_Rfuncrtions.R')
# MissPattern = GL_flash(Boolean_count,K = 10)
# saveRDS(MissPattern,file = "~/HG/LogisticFlash/data/MissPatternMouseBuettnerESC_K10.rds")
MissPattern = readRDS("~/HG/LogisticFlash/data/MissPatternMouseBuettnerESC_K10.rds")
```

I get 9 factors, but I cut it to four based on the PVE...(need more careful at that).

```{r,echo=FALSE}
par(mfrow = c(2,2),mar = c(5,4,4,2) - 1.8)
for( i in 1:4) barplot(MissPattern$f[,i],main = paste("factor",i))
total_counts = apply(counts,2,sum)
for( i in 1:4) plot(total_counts,MissPattern$f[,i],main = paste("factor",i,"vs","total counts"))
```
## factor 1 for total missing

this is similar for all studies.

## factor 2 for the outlier 

we can see that factor are mainly tell use about the ourlier of the total count, even if we only use the binary data.

## factor 3 and factor 4 for the cell cycle

```{r}
phase_lable = colnames(counts)
which_G1 <- grep("G1", colnames(counts))
which_G2 <- grep("G2", colnames(counts))
which_S <- grep("S_", colnames(counts))
phase_lable[which_G1] = "G1"
phase_lable[which_S] = "S"
phase_lable[which_G2] = "G2"
library(ggplot2)
plot_df = data.frame(MissPattern$f)
colnames(plot_df) = c("f1","f2","f3","f4")
plot_df$total_count = total_counts
p <- ggplot(plot_df, aes(total_count, f3, label = (as.character(phase_lable) )))
# p + geom_text()
p + geom_label(aes(fill = factor(phase_lable)), colour = "white", fontface = "bold")
p <- ggplot(plot_df, aes(total_count, f4, label = (as.character(phase_lable) )))
# p + geom_text()
p + geom_label(aes(fill = factor(phase_lable)), colour = "white", fontface = "bold")
```

## Session Information

```{r session-info}
```
